Towards a Cognitive System That Can Recognize Spatial Regions Based on Context

### 2012-01-23 scenarios/cdsr/trunk/papers/AAAI2012/main-aaai12-cdsr.tex
hacked intro

% nah: These are the main claims from the ACS paper, but they can be reduced/dropped for aaai

Representing and reasoning about context-dependent spatial regions is beyond the capabilities of current cognitive systems, yet it is an important ability for three reasons:

\begin{itemize}

\item{If cognitive systems are to collaborate with humans in everyday environments, then they must be able to represent and reason about the spatial regions humans refer to. Many regions are best defined in a context-dependent manner, for example, a kitchen in a studio apartment, an aisle in a church or store, behind enemy lines in a military engagement, etc.}

\item{Cognitive systems must integrate different types of information, including geometric, semantic, and functional knowledge. The identification of context-dependent spatial regions provides a suitable task for studying this integration. Only through such integration could a system determine, for example, whether a banquet hall has a dancefloor depending on the arrangement of the room.}

\item{Cognitive systems must adapt their knowledge and experience to new situations. While humans do not demonstrate mastery after seeing a single example of a new concept, they are able to begin using it immediately, whilst incrementally refining it over time (cf. one-shot learning~\cite{Fei-Fei/etal:2006}). This has important implications with respect to the amount of knowledge engineering and training time required to create a cognitive system. For example, having trained a mobile robot to understand the front of one classroom, it is desirable for it to be able to identify the front of other classrooms and similar areas (e.g. theaters) without additional tutoring.}
\end{itemize}

@@

This paper presents a novel integration of technologies which allows an artificial cognitive system (specifically a mobile robot) to represent and reason about CDSRs. Our approach is founded on two assumptions. The first assumption is that CDSRs can be defined using qualitative spatial representations (QSRs) extracted from the sensor data of the system~\cite{Cohn:2001}.

The second assumption is that semantically similar areas (e.g. two different classrooms) will feature similar CDSRs, and that these similarities can be recognised through analogy.

The rest of the paper is structured following these assumptions. Section~\ref{sec:qsr-gen} describes how we generate QSRs from the spatial model of an existing, state-of-the-art, cognitive system and use these to define CDSRs. Section~\ref{sec:analogy} then describes how use the structure-mapping model of analogy \cite{Gentner1983a} to transfer a CDSR from a labelled example to a new situation. Section~\ref{sec:example} presents a worked example of the entire process, and Section~\ref{sec:evaluation} evaluates our system in comparison to data from human subjects performing the same task.

\section{From Sensors to Semantics}\label{sec:qsr-gen}

@@

\subsection{Representing Context-Dependent Spatial Regions}

@@

\begin{figure}[t]
  \includegraphics[width=\columnwidth]{images/worked-example-rooms.png}
  \caption{\textbf{TODO}}
  \label{fig:rooms}
\end{figure}

\begin{figure}[t]
  \includegraphics[width=\columnwidth]{images/ug40.png}
  \caption{\textbf{TODO}}
  \label{fig:ug40}
\end{figure}

\from{Nick}{I'd like to keep an example like this in the paper, but we need to choose carefully so we can minimise the number of space-consuming images we need}

@@

Typical approaches to spatial representation for mobile robots tend to focus on localization, and thus mostly represent the world uniformly without subdivision into meaningful (semantic) units \cite{Thrun02a}. When a more structured representation is required (e.g., for behavior planning or human-robot interaction), topological maps (i.e., graphs of connected regions) are usually built on top of these metric representations using various landmark-detection techniques to segment regions, e.g. door detection \cite{Hawes/etal:2009b} or clustering of perceptual features \cite{Peltason/etal:2009a}. While capable of recognizing regions, these approaches do not consider the semantic information necessary for identifying context-dependent spatial regions. While steps have been taken toward incorporating this knowledge  (i.e. the types of objects present and their arrangement) into spatial representations, current approaches fall short of representing the types of regions discussed in this work. For example, a number of systems attempt to recognize the category of a room by the objects present within it \cite{Hawes/etal:2010b}, and others are able to mediate their behavior when certain functional contexts are detected (e.g. the areas near doors \cite{Zender2008a}).

% Nick: Goal reasoning is a separate issue entirely, and one for a different paper
 % Additionally, existing systems which are able to reason about their behavior are not able to do so with reference to the spatial extent of their goals, and thus would also fail to generate the desired behavior on this count.

There is mounting evidence that analogy, operating over structured qualitative representations can be used to simulate a number of spatial reasoning tasks. Forbus \textit{et al.} showed that analogy between course of action diagrams could be used to identify potential ambush locations in new situations by focusing on only the relevant aspects of sketched battle plans \cite{Forbus/etal2003}. A core contribution of their work was the definition of a \textit{shared similarity constraint} between a spatial reasoning system and its user; where users and spatial reasoning systems agree on the similarities between situations. This has close parallels to what we are trying to accomplish, where a cognitive system is able to reason about context-dependent spatial regions by identifying the same salient features as its human user. The anchor points in our work were originally used in teaching a system how to solve problems from the Bennett Mechanical Comprehension Test that require spatial and conceptual reasoning. For example, identifying which wheelbarrow will be more difficult to lift based on the relative locations of its loads as depicted in a sketch \cite{Klenk/etal2005}. In that work, the anchor points defined the endpoints of lines. We go beyond that result to use anchor points to specify 2D regions.  The combination of analogy and qualitative representation has been shown to be flexible enough to model problems at different levels of refinement (e.g. edges, shapes, groups) \cite{Lovett&Forbus2011} and to combine semantic as well as geometric information in spatial reasoning tasks (e.g. \cite{Lockwood/etal2008}). Our work demonstrates that the combination of analogy and qualitative spatial representations is useful for problems in cognitive mobile robotics. 
% Therefore, we hope and expect to see more research combining these three fields.

>>>>

Representing and reasoning about CDSRs is beyond the capabilities of current cognitive systems, yet it is an important ability. If cognitive systems are to collaborate with humans in everyday environments then they must be able to represent and reason about the spatial regions humans refer to. Many regions are best defined in a context-dependent manner, for example, a kitchen in a studio apartment, an aisle in a church or store, behind enemy lines in a military engagement, etc. In order to represent and reason about such regions, cognitive systems must integrate different types of information, including geometric, semantic, and functional knowledge. Thus, context-dependent spatial regions provide a suitable context for studying this integration.

% Cognitive systems must adapt their knowledge and experience to new situations. While humans do not demonstrate mastery after seeing a single example of a new concept, they are able to begin using it immediately, whilst incrementally refining it over time (cf. one-shot learning~\cite{Fei-Fei/etal:2006}). This has important implications with respect to the amount of knowledge engineering and training time required to create a cognitive system. For example, having trained a mobile robot to understand the front of one classroom, it is desirable for it to be able to identify the front of other classrooms and similar areas (e.g. theaters) without additional tutoring.

@@

This paper presents a novel integration of technologies which allows an artificial cognitive system (specifically a mobile robot) to represent and reason about CDSRs. Our approach is founded on two assumptions. The first assumption is that CDSRs can be defined using qualitative spatial representations (QSRs) extracted from the sensor data of the system~\cite{Cohn:2001}. The second assumption is that semantically similar areas (e.g. two different classrooms) will feature similar CDSRs, and that these similarities can be recognised through analogy. The rest of the paper is structured following these assumptions. Section~\ref{sec:qsr-gen} describes how we generate QSRs from sensor data taken from an existing, state-of-the-art, cognitive system and use these to define CDSRs. Section~\ref{sec:analogy} then describes how use the structure-mapping model of analogy \cite{Gentner1983a} to transfer a CDSR from a labelled example to a new situation. Section~\ref{sec:example} presents a worked example of the entire process, and Section~\ref{sec:evaluation} evaluates our system in comparison to data from human subjects performing the same task.

\section{Metric to Qualitative Representations}\label{sec:qsr-gen}

@@

\subsection{Representing CDSRs}

@@

% all images in a single figure -- pretty ugly and a waste of space
% \begin{figure*}[t]
%   \includegraphics[width=\textwidth]{images/worked-example.png}
%   \caption{\textbf{TODO}}
%   \label{fig:rooms-and-pic}
% \end{figure*}

\begin{figure}
  \includegraphics[width=\columnwidth]{images/worked-example-rooms.png}
  \caption{Maps of 2 real classrooms generated by our system, plus an expert-annotated CDSR (orange, top image), subject-annotated CDSR (red, bottom image) and a CDSR transferred by analogy (green bottom image). The classroom used to generate the top classroom is pictured in Figure~\ref{fig:ug40}}
  \label{fig:rooms}
\end{figure}

\begin{figure}
  \includegraphics[width=\columnwidth]{images/ug40.png}
  \caption{One of the classrooms used in our evaluation. This image was presented to subjects who were  asked to annotate a copy of the image in the bottom half of Figure~\ref{fig:rooms}.}
  \label{fig:ug40}
\end{figure}

@@

% robots and space

Typical approaches to spatial representation for mobile robots tend to focus on localization, and thus mostly represent the world uniformly without subdivision into meaningful (semantic) units \cite{Thrun02a}. When a more structured representation is required, many turn to Kuipers' Spatial Semantic Hierarchy~\cite{Kuipers:2000}. This paper follows in this tradition, adding CDSRs to his qualitative topological representations. Whilst mobile robots exist which can determine the type of a room from the objects in it ~\cite{Hanheide/etal:2010a,Galindo/etal:2005a}, they only concern themselves with types of whole rooms, and cannot represent regions within rooms. This is also true for those systems which use some elements of QSR~\cite{aydemir2011icra}. The need for an autonomous system to ground references to human-generated descriptions of space has been recognised in domains where a robot must be instructed to perform a particular task, however existing systems are restricted to purely geometrically-defined regions~\cite{Tellex:2011,Dzifcak/etal:2009,brenneretal07ijcai}.

There is mounting evidence that analogy, operating over structured qualitative representations, can be used to simulate a number of spatial reasoning tasks. Forbus \textit{et al.} showed that analogy between course of action diagrams could be used to identify potential ambush locations in new situations by focusing on only the relevant aspects of sketched battle plans \cite{Forbus/etal2003}. A core contribution of their work was the definition of a \textit{shared similarity constraint} between a spatial reasoning system and its user; where users and spatial reasoning systems agree on the similarities between situations. This has close parallels to what we are trying to accomplish, where a cognitive system is able to reason about context-dependent spatial regions by identifying the same salient features as its human user. The anchor points in our work were originally used in teaching a system how to solve problems from the Bennett Mechanical Comprehension Test that require spatial and conceptual reasoning. For example, identifying which wheelbarrow will be more difficult to lift based on the relative locations of its loads as depicted in a sketch \cite{Klenk/etal2005}. In that work, the anchor points defined the endpoints of lines. We go beyond that result to use anchor points to specify 2D regions.

% The combination of analogy and qualitative representation has been shown to be flexible enough to model problems at different levels of refinement (e.g. edges, shapes, groups) \cite{Lovett&Forbus2011} and to combine semantic as well as geometric information in spatial reasoning tasks (e.g. \cite{Lockwood/etal2008}). Our work demonstrates that the combination of analogy and qualitative spatial representations is useful for problems in cognitive mobile robotics. 
% Therefore, we hope and expect to see more research combining these three fields.
